---
title: "Initial abundances"
author: "Maude Vernet"
date: "2025-05-20"
output: html_document
---

# Monitoring data:
* monitoring patches in Palmyra are 100m2
* Palmyra size is 1190 ha -> 1.19*10^5 100m2
* preys taken into account if >25-180 mm 
* round selected for each species:
** highest recorded median abundance/patch
* lower and upper bounds based on 10 and 90% quantiles


```{r means per sampling sessions}
rm(mean_per_session)
for (i in 1:5) {
  round_processed_rdata_file <- paste0("data Sabrina/processed_data_round_", i, ".RData")
  load(round_processed_rdata_file)
  rm(taxa_data_percent, taxa_data_summary)
  
  # 1) patch means
  patch_means <- aggregate(Number ~ group + patch, 
                           data = taxa_data_long, FUN = mean)
  
  # 2) overall + CI by hand
  group_list <- split(patch_means$Number, patch_means$group)
  group_summary <- do.call(rbind, lapply(names(group_list), function(g) {
    x <- group_list[[g]]
    m <- mean(x)
    s <- sd(x) / sqrt(length(x))
    ci <- qt(0.975, df = length(x) - 1) * s
    data.frame(session = i,
               group = g,
               mean = m * 1.19 * 10^5)
  }))
  
  if (!exists("mean_per_session")) {
    mean_per_session <- group_summary
  } else {
    mean_per_session <- rbind(mean_per_session, group_summary)
  }
  rm(group_summary, taxa_data_long, group_list, patch_means)
}

mean_per_session
```

```{r extract data}
group_list <- split(mean_per_session$mean, mean_per_session$group)
Data_allround <- do.call(rbind, lapply(names(group_list), function(g) {
  x <- group_list[[g]]
  m <- mean(x)
  s <- sd(x) / sqrt(length(x))
  ci <- qt(0.975, df = length(x) - 1) * s
  data.frame(group = g,
             mean = m,
             lower_95CI = m - ci,
             upper_95CI = m + ci,
             n_sessions = length(x))
}))

rm(group_list)

Data_allround
```

# Other data

### Seabirds
- sooty-tern: 139 734 (in 2002);
- black noddies: ~20 000
- 750 000 nests (in 1993)
- more than 1 million nesting seabirds
Source: Ramsar Information Sheet for Palmyra Atoll (USFWS 2001)
https://rsis.ramsar.org/RISapp/files/RISrep/US1971RIS.pdf?utm_

#### bounds:
- lower bound: 10^6/ whole island
- upper bound: 1.5*10^6/ whole island

```{r seabirds}
seabirds_lb <- 10^6
seabirds_ub <- 1.5*10^6
```

### herbivorous crabs
- Tokelau Atoll Land‐Crab Density
- 560 crabs per 5 000 m2 -> 0.112 crabs/m2 on a similar Pacific atoll.
- ~1.3 * 10^6 On Palmyra (1.19 * 10^7 m2)
Source: Smithsonian, 1979
https://repository.si.edu/bitstream/handle/10088/5094/00235.pdf?utm_

#### bounds:
- lower bound: 1.2*10^6/ whole island
- upper bound: 1.4*10^6/ whole island
```{r seabirds}
herbivorous_crabs_lb <- 1.2*10^6
herbivorous_crabs_ub <- 1.4*10^6
```

### Carnivorous crabs
- Aldabra Atoll Coconut Crab (Birgus latro) Surveys
- 23.7/ha for coconut crabs on Aldabra
- ~28'203 On Palmyra (1190 ha)
Source: Appoo et al., 2021
https://www.jstor.org/stable/27060587?utm_
#### bounds:
- lower bound: 20'000/ whole island
- upper bound: 38'000/ whole island

```{r seabirds}
carnivorous_crabs_lb <- 20000
carnivorous_crabs_ub <- 38000
```

### native trees
- Cocos nucifera 47% of island coverage in 2012
- Rest = ~53% = 630.7 ha
- typical tropical-forest tree density = ≈ 1000 – 3000 trees/ha
- 630'700 - 1'892'100
Source:  Fact‐Sheet on Palmyra Atoll (USGS 2012)
https://pubs.usgs.gov/fs/2012/3002/pdf/fs20123002.pdf?utm_

#### bounds:
- lower bound: 630'700/ whole island
- upper bound: 1'892'100/whole island

```{r native trees}
native_trees_lb <- 630700
native_trees_ub <- 1892100
```

### palm trees
- 32'876 mature trees across the atoll (Struckhoff 2019)
- 31'839 adult trees removed from 2019 to present (319 weeks, 73.5 months) (3'309/ 4 months in between Feb and June 2025)
Source = Wolf et al. 2018
https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0200743&type=printable

```{r palm trees}
palm_trees_lb <- 30000
palm_trees_ub <- 65000
```

```{r combine data}
pre_lb <- round(
  c(palm_trees_lb,
    seabirds_lb,
    herbivorous_crabs_lb,
    carnivorous_crabs_lb,
    Data_allround[Data_allround$group == "cane_spiders",]$lower_95CI,
    Data_allround[Data_allround$group == "geckos",]$lower_95CI,
    Data_allround[Data_allround$group == "cockroaches",]$lower_95CI,
    Data_allround[Data_allround$group == "terrestrial_arthropods",]$lower_95CI,
    Data_allround[Data_allround$group == "flying_arthropods",]$lower_95CI,
    native_trees_lb,
    Data_allround[Data_allround$group == "others",]$lower_95CI))

save(pre_lb, file = "pre_lb.RData")

pre_ub <- round(
  c(palm_trees_ub,
    seabirds_ub,
    herbivorous_crabs_ub,
    carnivorous_crabs_ub,
    Data_allround[Data_allround$group == "cane_spiders",]$upper_95CI,
    Data_allround[Data_allround$group == "geckos",]$upper_95CI,
    Data_allround[Data_allround$group == "cockroaches",]$upper_95CI,
    Data_allround[Data_allround$group == "terrestrial_arthropods",]$upper_95CI,
    Data_allround[Data_allround$group == "flying_arthropods",]$upper_95CI,
    native_trees_ub,
    Data_allround[Data_allround$group == "others",]$upper_95CI))

save(pre_ub, file = "pre_ub.RData")
```
